### 2011-04-07 papers/phase-one/paper.tex
Add sam's new section, and section on security.

\documentclass{dependencies/acm_proc_article-sp}

\usepackage{url}
\usepackage{color}
\usepackage{verbatim}
\usepackage{listings}
\lstset{
  language=C++,             % choose the language of the code
  basicstyle=\small,       % the size of the fonts that are used for the code
  numbers=left,                   % where to put the line-numbers
  numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
  stepnumber=0,                   % the step between two line-numbers. If it is 1 each line will be numbered
  numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  %frame=single,                   % adds a frame around the code
  tabsize=2,              % sets default tabsize to 2 spaces
  captionpos=t,                   % sets the caption-position to bottom
  breaklines=false,        % sets automatic line breaking
  breakatwhitespace=false,    % sets if automatic breaks should only happen at whitespace
  escapeinside={\%}{)}          % if you want to add a comment within your code
}

% Get rid of the permission block
\makeatletter
\let\@copyrightspace\relax
\makeatother

\begin{document}

%\title{ Distributed Systems: Project Description }
\title{ Distrivia: A Distributed Trivia Game }
\numberofauthors{3}
\author{
\alignauthor
Brian Gianforcaro \\
       \affaddr{Rochester Institue of Technology}\\
       \email{bjg1955@rit.edu}
\alignauthor
Steven Glazer \\
       \affaddr{Rochester Institue of Technology}\\
       \email{sfg6126@rit.edu}
\alignauthor
Samuel Milton \\
       \affaddr{Rochester Institue of Technology}\\
       \email{srm2997@rit.edu}
}
\maketitle

%\begin{abstract}
%In this paper we detail our initial idea for a distributed trivia game.
%We will explain general game play idea's as well as an example of the
%possible architecture for our system.
%\end{abstract}

\section {What We've Done}
So far, our work has mainly focused on getting everything setup to work as a
distributed system. Getting the ground work done (for example servers and load
distributers running) has provided us a way to quickly develop the clients
and server as well as provided us with a way to reliably connect the servers to
the clients.
\subsection {Servers}
Currently, we have 4 servers running. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load distributers. This will help us manage if the services
are running and allow us to track down any problems, if they arise. The other three
servers are all running our server application and hosting the databases.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 3 servers. If
we were to add a server, we would just need to update Riak and it would sync the
databases to any additional servers as well. These servers, therefore, all contain
the same information. If any servers go down, the distributed system would still
run fine. We could have all servers go down except one and still be reliably
handling requests.
\subsection {Traffic}
We have two load distributers set up. Both distributers are currently configured to
divide the requests among the 3 active servers. The duplicate distributers allow
for a load distributer to go down and we would still be able to serve requests
to the servers. These load distributers are selected by the clients using round
robin DNS where distrivia.lame.ws will select either load distributer that is
up at random. This provides us will stability for access to the servers. The
load distributers are set up the attempt to connect to a server twice before
deeming it dead and removing it from its list of servers to direct clients towards.
\subsection {Messages}
We are currently using https protocol for our messages. This provides a secure
and usable message system that we have conformed our client and server APIs to use.
The actual message content is in the form of JSON objections \cite{json} 
\subsection {Clients}
Our focus for clients has been on a web-based client and an android application.
Both clients have most of the their screens completed. The web-based client will
be run from desktop platforms and laptop platforms. The android client will run
on any android phone and does not use any unique services that require specific
hardware. The web client is nearly fully-functional. It currently can login/register
players, join public games, and compete in public games. The android client has
he UI done for all the different activities. The android client successfully registers
and logs in players.

\section {What's Left}
\subsection {Web Client}
 Joining/playing private games
 In-game leaderboard
 inish connecting to actual server API
 Implement useful user warnings
\subsection {Android}
So far, the only connection to the server that the android client has is when it
registers/logs in. We still need to connect the rest of the android client to
the server API to make sure all the communication between the android client
and the server are working. This includes joining public/private rounds,
competing in these rounds, and viewing leaderboards. While the UI is nearly
finished on the android client, nearly all of the connections to the servers
has yet to be done. Additionally, the android client needs to have more user
feedback and increased ability for the users to control which activity they
are in.
\subsection {Server}
Private game joining
Global leaderboard

\section {Meeting the Requirements}

Our system has two unique clients. One of them is designed for the Android platform as an app.
The other is a web client which is specifically being tested on Desktop and Laptop
browsers, however it may work on many handheld web browsers as well.

Our system supports having any number of servers and allowing any subset of them to fail.  We are using Riak, which supports a ''shared nothing' architecture.  We can add a new server by joining it to our database and starting the web server.  This can be done as many times as needed, allowing us to scale up our system.  These same principles also give us great fault tolerance, since any server is capable of running independently.  As long as one server is still up, the service will be available.

We use round robin DNS to allow our clients to be completely ignorant of our servers.  Load distributors can be added and removed from the DNS server and the clients will be completely ignorant of this and unaffected.  The load distributors in turn know about our data servers and do their best to distribute the load across them.  When a new server is added, it is just added to our load distributor configuration file and clients can start using it.  When a server fails, the distributors will stop directing traffic to it.

- Coordination/Cooperation (Synchronous Databases and Rounds consisting of many players)
- Replication / Availability / Consistency (Riak handles synchronizing databases among servers. Load distributers provide availability)
  < Needs stuff on Riak>

 Our team is taking a very proactive stance towards security. All of our machines are properly fire walled so that
 outside clients can not authenticate to our database cluster. Also we have implemented SSL through our entire infrastructure.
 Our load balancer's automatically redirect http:// to https:// and once you are in https:// at the load balancer, SSL continues all the
 way back to our actual web application back end serving responses over a SSL wrapped socket. This way no passwords are ever
 sent in clear text. Our group even found a SSL certificate authority who was willing to sign our SSL certificate for free. This gives
 us actual fully functional, authenticate able SSL infrastructure. In order to make sure that our system properly secures user information,
 we stuck with the industry standard of storing a hashed version of the users password instead of the password itself. We decided on using
 the cryptographically secure hashing function designed for passwords known as bcyrpt \cite{bcrypt}.

- Use of an interesting distributed algorithm

\newpage
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{bibliography}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
\balancecolumns
% That's all folks!
\end{document}